---
title: "Employment in the tech sector"
output: html_document 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This document represents a brief insight of the Tech sector employment trends. The data I have used is publicly available at labor market information portal. 

Labor market portal has data related to different industries. I have picked up information media and telecommunications to get some insights of the tech sector.

```{r }
data_url <- "https://lmip.gov.au/PortalFile.axd?FieldID=3195450&.xlsx"
temp = tempfile(fileext = ".xlsx")
download.file(data_url, destfile=temp, mode='wb')
test <- readxl::read_excel(temp, sheet =2)

# extracting Tech data from table
tech_data <- test[test$Industry=="Information Media and Telecommunications",c(1,2,4,5,6,7,8,9,10 )]
```

#### Five year growth in tech industry w.r.t state/territory 
I have created a plot which shows five year growth in tech sector in each state and overall total growth in Australia.


```{r Five year growth in tech industry w.r.t state/territory }

tech_industry_Growth <- aggregate(tech_data$`Five year growth by Industry` ~ tech_data$`State/Territory`,FUN = sum)
colnames(tech_industry_Growth) <- c("State" , "Growth")
tech_industry_Growth

tech_industry_Growth %>%
  ggplot(aes(x = State, y=Growth)) +
           geom_col(aes(fill = State))+ 
           labs(title = "Five year growth in tech industry", x="State",y="Five year growth")
          

```

If we look at the graph of five years growth we can clearly see that except VIC, NT and NSW/ACT all states show decline in growth. NSW and QLD show largest decline among the states.

#### Male and Female in Tech Industry

I have created another plot which shows number of male and female working in tech sector in each state and overall Australia.


```{r Male and Female in Tech Industry}

tech_people <- aggregate(cbind(tech_data$`Employed - Male`,tech_data$`Employed - Female`) ~ tech_data$`State/Territory`,FUN = sum)
colnames(tech_people)<-c("State","Male","Female")
tech_people

tech_people %>%
  ggplot(aes(x = State)) +
      geom_col(aes(y = Male), width = 0.25,fill="blue",position = position_nudge(x =-0.125))+
      geom_col(aes(y = Female ) ,width = 0.25, fill="red",position =position_nudge(x=0.125))+          labs(title = "Male and Female in Tech Industry",x="States",y="total number")
```

Blue represents male population while red represents female  population.

